---
title: "Statistical Inference Course - ToothGrowth Data Analysis"
author: "Vijay Vaidya"
date: "Thursday, July 23, 2015"
output: pdf_document
---
##Overview
We will analyze the "ToothGrowth" data from R package and compare the impact of the toothgrowh with the supplement and dose of supplement

##Exploratory data analysis
We will load "ToothGrowth" data and do some basic analysis
```{r load, echo=TRUE}
rm(list = ls()) # clear the environment
data(ToothGrowth) # load the data
class(ToothGrowth) # check the class/data type
```

Now that we know that we have frame, we can do basic analysis
```{r, echo=TRUE}
#Total no.of rows
nrow(ToothGrowth)

#Column names
colnames(ToothGrowth)

#Supplement types
levels(ToothGrowth$supp)

#How many unique doses
unique(ToothGrowth$dose)

#Max length
ToothGrowth[ which.max(ToothGrowth$len), ]

#Min Length
ToothGrowth[ which.min(ToothGrowth$len), ]

#Avg Length
mean(ToothGrowth$len)

#Avg Length of tooth from the group taking OJ
mean(ToothGrowth[ ToothGrowth$supp == "OJ", ]$len)

#Avg Length of tooth from the group taking VC
mean(ToothGrowth[ ToothGrowth$supp == "VC", ]$len)

#Any NA values?
which(is.na(ToothGrowth$len))

```
we can now chart a graph to show consumption of orange juice or vitamin C and the doses impacts teeth growth:
```{r, echo=TRUE}
library(ggplot2)
require(graphics)
coplot(len ~ dose | supp, data = ToothGrowth, panel = panel.smooth,xlab = "ToothGrowth data: length vs dose, given type of supplement")
```

##Comparison of tooth growth by supp and dose
So we want to compare whether length of the tooth dependins upon the dose of OJ or VC.  So we split *ToothGrowth* by doses and calculate the respective confidence intervals:  
```{r, echo=TRUE}
doses <- unique(ToothGrowth$dose)
```

** Dose `r doses[ 1 ]` mg**

Avg tooth length with OJ - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 1 ] & ToothGrowth$supp == "OJ") , ]$len)` 
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 1 ] & ToothGrowth$supp == "OJ") , ]$len)`


Avg tooth length with VC - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 1 ] & ToothGrowth$supp == "VC") , ]$len)`
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 1 ] & ToothGrowth$supp == "VC") , ]$len)`

```{r, echo=TRUE}
d <- ToothGrowth[ ToothGrowth$dose == doses[ 1 ], ]
t.test(len ~ supp, data = d, paired = FALSE, var.equal = TRUE)
```

** Dose `r doses[ 2 ]`mg**

Avg tooth length with OJ - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 2 ] & ToothGrowth$supp == "OJ") , ]$len)` 
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 2 ] & ToothGrowth$supp == "OJ") , ]$len)`


Avg tooth length with VC - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 2 ] & ToothGrowth$supp == "VC") , ]$len)`
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 2 ] & ToothGrowth$supp == "VC") , ]$len)`

```{r, echo=TRUE}
d <- ToothGrowth[ ToothGrowth$dose == doses[ 2 ], ]
t.test(len ~ supp, data = d, paired = FALSE, var.equal = TRUE)
```

** Dose `r doses[ 3 ]`mg**

Avg tooth length with OJ - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 3 ] & ToothGrowth$supp == "OJ") , ]$len)` 
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 3 ] & ToothGrowth$supp == "OJ") , ]$len)`


Avg tooth length with VC - `r mean(ToothGrowth[ (ToothGrowth$dose == doses[ 3 ] & ToothGrowth$supp == "VC") , ]$len)`
Std deviation `r sd(ToothGrowth[ (ToothGrowth$dose == doses[ 3 ] & ToothGrowth$supp == "VC") , ]$len)`

```{r, echo=TRUE}
d <- ToothGrowth[ ToothGrowth$dose == doses[ 3 ], ]
t.test(len ~ supp, data = d, paired = FALSE, var.equal = TRUE)
```

##Conclusion
1. From the results we can conclude that when dose size s of 0.5mg, average tooth growth is higher when suppeent of OJ is provied than the supplement of VC. We can say that with 95% confidence because p-value (0.005304) is < 0.01. 
2. Same thing (as described above) is true for the dose size of 1mg. We have the strong evidence for thes dose size of 1mg because p-value (0.0007807) is too small.
3. However we can't cconclude the same for the dose of 2mg because p-value (0.9637) is way higher that 0.01
