---
title: "Statistical Inference Course Project - Simulation Exercise"
author: "Vijay Vaidya"
date: "Wednesday, July 22, 2015"
output: pdf_document
---
##Overview
Investigate exponential distribution in R and compare with Central Limit Theorem. Illustrate via simulation that mean and variance approach the theoratical values for large number of simulations. We will take 1000 simulations.



```{r setup, echo=TRUE}
rm(list = ls()) # clear the environment 
lambda <- 0.2   # set the rate parameter 
n <- 40         # exponentials' count
simCount <- 1000 # simulations' count
```

##Compare Sample mean with theoratical mean. 
We are given theoratical mean as `r 1 / lambda` (1/`r lambda`). We will create a vector "means" of `r simCount` elements. Each element is the mean of `r n` exponentials. We take the average of "means" and then plot the chart.

```{r domean, echo=TRUE}
theoreticalMean <- 1 / lambda
means <- c()
for(i in 1:simCount)
{
    means <- c(means, mean(rexp(n, lambda)))
}
m <- cumsum(means) / (1:simCount)

plot(x = 1:simCount,
     y = m,
     type = "l",
     main = "The sample mean versus the theoretical mean",
     xlab = "Simulations",
     ylab = "Mean",
     col = "black")

lines(x = c(1, simCount),
      y = c(theoreticalMean, theoreticalMean),
      type = "l",
      col = "blue")

legend("topright", pch = "_",
       col = c("black", "blue"),
       legend = c("sample mean", "theoretical mean"))
```

##Compare Sample variance with theoratical variance. 
We have theoratical variance as `r 1 / lambda^2` (1/`r lambda^2`)
will create a vector "var" of `r simCount` elements. Each element is the variance of `r n` exponentials. We take the average of "vars" and then plot the chart

```{r dovar, echo=TRUE}
theoreticalVar <- 1 / lambda^2
vars <- c()
for(i in 1:simCount)
{
    vars <- c(vars, var(rexp(n, lambda)))
}
v <- cumsum(vars) / (1:simCount)

plot(x = 1:simCount,
     y = v,
     type = "l",
     main = "The sample variance versus the theoretical variance",
     xlab = "Simulations",
     ylab = "variance",
     col = "black")

lines(x = c(1, simCount),
      y = c(theoreticalVar, theoreticalVar),
      type = "l",
      col = "blue")

legend("bottomright", pch = "_",
       col = c("black", "blue"),
       legend = c("sample variance", "theoretical variance"))
```

##Distribution.
Now we prove that as the sample size increases, distribution becomes standard  normal.
```{r dist, echo=TRUE}
hist(means,
     main = paste("Distribution of", simCount, "means"),
     xlab = "Mean")
```
